Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-07-31 manuscript.tex
added some plots of results and discussion.

	&= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}	\right)

@@

Applying a discrete version of the Lagrange-d?Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.

@@

A numerical test case is presented to demonstrate this process and generate preliminary results.
The goal is to determine a transfer trajectory between two planar periodic orbits of the Earth-Moon system. 
The initial and target orbits are depicted in Figure~\ref{fig:opt_reach_trajectory} as the black periodic orbits.
From the initial state, a series of optimal trajectories are generated which all intersect the desired Poincar\'e section.
The intersection of these trajectories with the Poincar\'e section is shown in Figure~\ref{fig:north_poincare_reach}.
The reachable set is discretely approximated as the area enclosed by the optimal trajectories.
As the target state does not lie within the reachable set, another iteration of the optimal control calculation must be performed. 
The location on the reachable surface which lies closest to the target state is used to generate the initial conditions for the next iteration.
A second, or southern, trajectory calculation once again generates an approximation of the reachable set as shown in Figure~\ref{fig:south_reach}.
In this iteration, that target state is enclosed within the reachable set. 
\Cref{fig:opt_reach_trajectory} illustrates the completed optimal transfer trajectory.
As the solution is composed of two distinct optimal solutions there exists a discontinuity at the switching point.
This discontinuity is most evident in the control history shown in \Cref{fig:opt_reach_control}.
More general transfers would require multiple applications of this reachability computation and additional Poincar\'e sections.
This recursive method allows one to enlarge the attainable states on the Poincar\'e section.
General transfers are possible by combining powered segments along the invariant manifolds.

Maximize a distance on the Poincar\`e section

Necessary conditions for optimality

Discrete version of costate equations of motion

Avoid explicit inverse by using gauss jordan elimination

Plot to compare continuous and discrete costate equations of motion

\section{Numerical Example}\label{sec:simulation}
Transfer between a periodic orbit and the Moon

Multiple Shooting method (choose interior patch points to decrease sensitivity)

Determine reachable set on Poincar\`e Section

Intersection of the sets on section

Minimization to a target state

Show transfer to the moon

Compare with the Ross work and the invariant manifold ( long time of flight)
\section{Conclusion}\label{sec:conclusion}

>>>>

	&= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}	\right) \, dt

@@

Applying a discrete version of the Lagrange-d'Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.

@@

{\Large DISCUSS MULTIPLE SHOOTING}

\section{Numerical Example}\label{sec:simulation}
A numerical simulation is presented to demonstrate the transfer procedure.
The goal is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
The initial state is propagated over a long period of \( t = \num{20} \) in non dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the desired trajectories remain in the vicinity of the moon in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.

A periodic orbit is generated about the \( L_1 \) Lagrange point. 
A Poincar\`e section is chosed to allow for design on a lower dimensional subspace.
The section is defined along the \( x \) axis and intersects both the initial and target orbits.
From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
By varying the angle \( \theta\) a different direction along the Poincar\`e section is maximized. 
The intersection of the optimal trajectories as well as those of the target Moon orbit with the Poincar\`e section are shown in~\cref{fig:transfer_orbit}
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Controlled Trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{manifold_trajectory} 
		\caption{Invariant Manifold} \label{fig:manifold_trajectory} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
	\end{subfigure} 
	\caption{Transfer Trajectory}
	\label{fig:transfer_orbit} 
\end{figure}
The optimal controlled trajectories from~\cref{fig:reach_trajectory} generate the reachable set denoted by the red markers in~\cref{fig:poincare_compare}.
The blue points in~\cref{fig:poincare_compare} are the intersections of the target Moon orbit and the Poincar\`e section.
As the reachable set intersects the target orbit a transfer is achievable.
The intersection point is determined via a linear interpolation and a subsequent transfer trajectory is generated as shown by the green trajectory in~\cref{fig:reach_trajectory}.

As a comparison the method introduced in Reference~\citenum{koon2000} is implemented.
The method is based on the invariant manifolds associated with the periodic orbits of the three body system.
These invariant manifolds are a set of trajectories that either asymptotically arrive or depart the periodic orbit. 
Linking the invariant manifolds of several periodic orbits, or coupled three body systems, has been used to generate orbital transfers.
\Cref{fig:manifold_trajectory} shows the unstable invariant manifold associated with the initial periodic orbit. 
The intersections on the Poincar\'e section show that a portion of the invariant manifold fails to intersect with the target orbit.
In addition, the sections of the manifold that do intersect the target orbit have a much longer time of flight. 
On average, the time along the invariant manifold is roughly twice that of the trajectory found via the reachable set computation.
In addition, while the there exists an intersection on the Poincar\`e section there is no guarantee of all states intersecting.
As a result, another \( \Delta V \) is required to transfer from the invariant manifold to the target orbit.

%Avoid explicit inverse by using gauss jordan elimination

%Plot to compare continuous and discrete costate equations of motion

%Multiple Shooting method (choose interior patch points to decrease sensitivity)
%
%Determine reachable set on Poincar\`e Section
%
%Intersection of the sets on section
%
%Minimization to a target state
%
%Show transfer to the moon
%
%Compare with the Ross work and the invariant manifold ( long time of flight)
\section{Conclusion}\label{sec:conclusion}
